---
output:
  html_document: default
  pdf_document: default
---
# Insert chapter 2 title here

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods.
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using.

#Konrad Sopyllo

###Data wrangling

#In this Data wrangling part we will focus brielfy on creating the data set as in the csv-file in the data-folder. 

```{r}

library(tidyverse)

url<-"http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS3-data.txt"
data=read_tsv(url)

```

#Creating an analysis_dataset with the variables gender, age, attitude, deep, stra, surf and points 
#by combining questions in the learning2014 data.

```{r}
collumns <- c("gender", "Age", "Attitude", "Points")

analysis_dataset <- data[collumns]
```

#Stra =  st_os+st_tm = ST01+ST09+ST17+ST25+ST04+ST12+ST20+ST28
```{r}
stra_collumns = unlist(str_split(c("ST01+ST09+ST17+ST25+ST04+ST12+ST20+ST28"), "\\+"))
```

#surf = su_lp+su_um+su_sb = SU02+SU10+SU18+SU26 + SU05+SU13+SU21+SU29 + SU08+SU16+SU24+SU32
```{r}
surf_collumns = unlist(str_split(c("SU02+SU10+SU18+SU26+SU05+SU13+SU21+SU29+SU08+SU16+SU24+SU32"), "\\+"))
```

#deep = d_sm+d_ri+d_ue = D03+D11+D19+D27 + D07+D14+D22+D30 + D06+D15+D23+D31
```{r}
deep_collumns = unlist(str_split(c("D03+D11+D19+D27+D07+D14+D22+D30+D06+D15+D23+D31"), "\\+"))
```


# creating scaled variables using rowMeans-function.
```{r}
analysis_dataset["stra"] = rowMeans(select(data, one_of(stra_collumns)))
analysis_dataset["deep"] = rowMeans(select(data, one_of(deep_collumns)))
analysis_dataset["surf"] = rowMeans(select(data, one_of(surf_collumns)))
```


# filtering out rows where Points variable = 0
```{r}
analysis_dataset = analysis_dataset[analysis_dataset$Points != 0,]
dim(analysis_dataset)

```

#as in the exercise provided the file has 166 observations and 7 variables

#Saving the analysis_dataset as a csv-file
```{r}

library(readr)
write_csv(analysis_dataset, "data/learning2014.csv")

```

#Showing how to read the csv-file

```{r}
loaded_data <- read_csv("data/learning2014.csv")
str(loaded_data)
head(loaded_data)
```

###Data analysis

The data presented is based on a survey and the data was collected 3.12.2014 - 10.1.2015 in Florence,Italy. The data was translated in Finnish by Kimmo Vehkalahti and Liisa Myyry. The survey was based originally on a finnish study concerning the progress and satisfaction related to the double phased study programme in technical universities. 

#correlation matrix

```{r}

library("Hmisc")
library("corrplot")
library(ggplot2)
```

#Gender correlation
```{r}
gender_graph <- ggplot(data.frame(analysis_dataset), aes(x = gender)) +
  geom_bar(fill = "lightgreen")
print(gender_graph)
```

#Points correlation
```{r}
points_dist <- ggplot(data.frame(analysis_dataset), aes(x =Points)) +
  geom_density(fill = "lightgreen")
print(points_dist)
```

#Age correlation
```{r}
age_dist <- points_dist <- ggplot(data.frame(analysis_dataset), aes(x = Age)) +
  geom_density(fill = "lightgreen")
print(points_dist)
```

#deep correlation
```{r}
deep_dist <- ggplot(data.frame(analysis_dataset), aes(x = deep)) +
  geom_density(fill = "lightgreen")
print(deep_dist)
```

#surf correlation
```{r}
surf_dist <- ggplot(data.frame(analysis_dataset), aes(x = surf)) +
  geom_density(fill = "lightgreen")
print(surf_dist)
```

#stra correlation
```{r}
stra_dist <- ggplot(data.frame(analysis_dataset), aes(x = stra)) +
  geom_density(fill = "lightgreen")
print(stra_dist)
```


#Creating a correlation graph of variables:"Attitude","deep","stra","surf","Points" using Pearson
```{r}
correlations = c("Attitude","deep","stra","surf","Points")
analysis_cor = cor(analysis_dataset[correlations])
corrplot(analysis_cor)


plot(analysis_dataset)

```

# By having a look at the correlation analysis we can identify the factors with the highest correspondance. Those will be further explained below.


##Attitude vs Points
```{r}
p1 <- ggplot(data.frame(analysis_dataset), aes(x = Attitude, y = Points))
p2 <- p1 + geom_point()
p3 <- p2 + geom_smooth(method = "lm")
p4 <- p3+ggtitle("Attitude vs Points")
print(p4)


model <- lm(analysis_dataset$Points~analysis_dataset$Attitude)
summary(model)
```

#p-value being lowest in this correlation graph

# Other way of doing the same thing
```{r}

library(ggplot2)
qplot(Attitude, Points, data = analysis_dataset) + geom_smooth(method = "lm")
my_model <- lm(Points ~ Attitude, data = analysis_dataset)

```

##surf vs deep 

```{r}
p1 <- ggplot(data.frame(analysis_dataset), aes(x = deep, y = surf))
p2 <- p1 + geom_point()
p3 <- p2 + geom_smooth(method = "lm")
p4 <- p3+ggtitle("deep vs surf")
print(p4)

model <- lm(analysis_dataset$surf~analysis_dataset$deep)
summary(model)

```
# p-value being second lowest in this comparison

##Attitude vs surf
```{r}
p1 <- ggplot(data.frame(analysis_dataset), aes(x = Attitude, y = surf))
p2 <- p1 + geom_point()
p3 <- p2 + geom_smooth(method = "lm")
p4 <- p3+ggtitle("Attitude vs surf")
print(p4)

model <- lm(analysis_dataset$surf~analysis_dataset$Attitude)
summary(model)
```
#p-value being third lowest in this comparison.

#multiple variable model
```{r}

library(GGally)
library(ggplot2)
ggpairs(analysis_dataset, lower = list(combo = wrap("facethist", bins = 20)))
my_model2 <- lm(Points ~ Attitude + stra, data = analysis_dataset)
summary(my_model2)


my_model2 <- lm(Points ~ Attitude + stra, data = analysis_dataset)
```
# Linear model with multiple variables
```{r}
lm(my_model2)

par(mfrow = c(2, 2))
```
# Residuals vs Fitted values (which = 1)
```{r}
plot(my_model2, which = 1)
```
# Normal QQ-plot (which = 2)
```{r}
plot(my_model2, which = 2)
```
# Residuals vs Leverage (which = 5)
```{r}
plot(my_model2, which = 5)
```

#Multiple explanatory variables all in one
```{r}
plot(my_model2, which = c(1, 2, 5))
```

#Making predictions

```{r}
m <- lm(Points ~ Attitude, data = analysis_dataset)
summary(m)
```
#New observations
```{r}
new_attitudes <- c("Mia" = 3.8, "Mike"= 4.4, "Riikka" = 2.2, "Pekka" = 2.9)
new_data <- data.frame(Attitude = new_attitudes)
print(new_data)
```

##Predictions
```{r}
predictions <- predict(m, newdata = new_data)
print(predictions)
```

#following the liner model the predicted values are Adjusted R-squared:  0.09939. This means the model doens not quite well predict the values.






```

Here we go again...
