Konrad Sopyllo

##Assignment 3 


The data presented below concers student performance in Portugese high school education.  It focuses mainly onperformance in two distinct subjects: Mathematics (mat) and Portuguese language (por). The data was collected by Cortez and Silva in 2008.
```{r}
url <- "https://raw.githubusercontent.com/KimmoVehkalahti/Helsinki-Open-Data-Science/master/datasets/alc.csv"
alc_data <- read.csv(url)

variable_names <- names(alc_data)
cat("Variable Names:", paste(variable_names, collapse = ", "), "\n\n")

```
In this section we will choose four random factors and analyse their correlation with alcohol consumption. First we will make a hypothesis on the result and then calculate it. In addition we will interpret coefficients of the models and their odds ratios

#Hypothesis 1. Romantic relationship is correlated with higher alcohol use.

```{r}
library(tidyverse)
gathered_data <- alc_data %>%
  gather(key = "variable", value = "alcohol_consumption", Dalc, Walc)

ggplot(gathered_data, aes(x = romantic, fill = variable, y = alcohol_consumption)) +
  geom_bar(position = "dodge", stat = "summary", fun = "mean", color = "black", alpha = 0.7) +
  labs(title = "Average Alcohol Consumption vs Romantic relationship",
       x = "romantic",
       y = "Average Alcohol Consumption") +
  scale_fill_manual(values = c("red", "green"), name = "Alcohol Consumption",
                    labels = c("Workday", "Weekend")) +
  theme_minimal()

logistic_model <- glm(high_use ~ Dalc + Walc + romantic, 
                      data = alc_data, 
                      family = binomial)

summary(logistic_model)

odds_ratios <- exp(coef(logistic_model))
conf_intervals <- exp(confint(logistic_model))

cat("\nOdds Ratios and 95% Confidence Intervals:\n")
data.frame(Odds_Ratio = odds_ratios, 
           Lower_CI = conf_intervals[, 1], 
           Upper_CI = conf_intervals[, 2])

```
Romantic relationship decreases the chances of high alcohol use. The odds ratio is 0.8 which means that the chances are smaller when in romantic realtionship. 

#Hypothesis 2. more activities is related to higher alcohol use. 
```{r}
gathered_data <- alc_data %>%
  gather(key = "variable", value = "alcohol_consumption", Dalc, Walc)

ggplot(gathered_data, aes(x = activities, fill = variable, y = alcohol_consumption)) +
  geom_bar(position = "dodge", stat = "summary", fun = "mean", color = "black", alpha = 0.7) +
  labs(title = "Average Alcohol Consumption vs Activities",
       x = "Activities",
       y = "Average Alcohol Consumption") +
  scale_fill_manual(values = c("red", "green"), name = "Alcohol Consumption",
                    labels = c("Workday", "Weekend")) +
  theme_minimal()

#regression analysis

logistic_model <- glm(high_use ~ Dalc + Walc + activities, 
                      data = alc_data, 
                      family = binomial)

summary(logistic_model)

odds_ratios <- exp(coef(logistic_model))
conf_intervals <- exp(confint(logistic_model))

cat("\nOdds Ratios and 95% Confidence Intervals:\n")
data.frame(Odds_Ratio = odds_ratios, 
           Lower_CI = conf_intervals[, 1], 
           Upper_CI = conf_intervals[, 2])

logistic_model <- glm(high_use ~ Dalc + Walc + goout, 
                      data = alc_data, 
                      family = binomial)

summary(logistic_model)

odds_ratios <- exp(coef(logistic_model))
conf_intervals <- exp(confint(logistic_model))

cat("\nOdds Ratios and 95% Confidence Intervals:\n")
data.frame(Odds_Ratio = odds_ratios, 
           Lower_CI = conf_intervals[, 1], 
           Upper_CI = conf_intervals[, 2])
```
Extracurricular activities decrease alchol consumption.
-p-value is low thus the result is significant. 


#Hypothesis 3. Amount of going out is correlated with higher alcohol use

```{r}
gathered_data <- alc_data %>%
  gather(key = "variable", value = "alcohol_consumption", Dalc, Walc)

ggplot(gathered_data, aes(x = goout, fill = variable, y = alcohol_consumption)) +
  geom_bar(position = "dodge", stat = "summary", fun = "mean", color = "black", alpha = 0.7) +
  labs(title = "Average Alcohol Consumption vs Going out",
       x = "Going out",
       y = "Average Alcohol Consumption") +
  scale_fill_manual(values = c("red", "green"), name = "Alcohol Consumption",
                    labels = c("Workday", "Weekend")) +
  theme_minimal()

logistic_model <- glm(high_use ~ Dalc + Walc + goout, 
                      data = alc_data, 
                      family = binomial)

summary(logistic_model)

odds_ratios <- exp(coef(logistic_model))
conf_intervals <- exp(confint(logistic_model))

cat("\nOdds Ratios and 95% Confidence Intervals:\n")
data.frame(Odds_Ratio = odds_ratios, 
           Lower_CI = conf_intervals[, 1], 
           Upper_CI = conf_intervals[, 2])

```
The more you go out, the more you drink alcohol. 
-Given the p-valuen (0.02) the results are statistically significant. 
-One unit increase in going out, alcohol consumption increases by 1.3


#Hypothesis 4. increased internet usage is correlated with smaller alcohol consumption.

```{r}
gathered_data <- alc_data %>%
  gather(key = "variable", value = "alcohol_consumption", Dalc, Walc)

ggplot(gathered_data, aes(x = internet, fill = variable, y = alcohol_consumption)) +
  geom_bar(position = "dodge", stat = "summary", fun = "mean", color = "black", alpha = 0.7) +
  labs(title = "Average Alcohol Consumption vs Internet Use",
       x = "Internet Use",
       y = "Average Alcohol Consumption") +
  scale_fill_manual(values = c("red", "green"), name = "Alcohol Consumption",
                    labels = c("Workday", "Weekend")) +
  theme_minimal()

logistic_model <- glm(high_use ~ Dalc + Walc + internet, 
                      data = alc_data, 
                      family = binomial)

summary(logistic_model)

odds_ratios <- exp(coef(logistic_model))
conf_intervals <- exp(confint(logistic_model))

cat("\nOdds Ratios and 95% Confidence Intervals:\n")
data.frame(Odds_Ratio = odds_ratios, 
           Lower_CI = conf_intervals[, 1], 
           Upper_CI = conf_intervals[, 2])


```
Internet use is increased during alcohol consumption. Thus the assumption is false.
-Given the p-valuen (0.02) the results are statistically significant. 
-One unit increase in internet usage, alcohol consumption increases by 1.2


#In the last section we do a prediction power model vs simple guessing of the results. 
```{r}
predictions <- predict(logistic_model, type = "response") > 0.5

conf_matrix <- table(Actual = alc_data$high_use, Predicted = predictions)

print(conf_matrix)

training_error <- sum(alc_data$high_use != predictions) / nrow(alc_data)
cat("\nTraining Error:", training_error, "\n")

library(ggplot2)
ggplot(alc_data, aes(x = as.factor(high_use), fill = as.factor(predictions))) +
  geom_bar(position = "dodge") +
  labs(title = "Actual vs. Predicted Values",
       x = "Actual Values",
       y = "Count",
       fill = "Predicted Values") +
  theme_minimal()

training_error <- sum(alc_data$high_use != predictions) / nrow(alc_data)
cat("\nTraining Error:", training_error, "\n")

majority_class <- as.numeric(names(sort(table(alc_data$high_use), decreasing = TRUE)[1]))
simple_guess_predictions <- rep(majority_class, nrow(alc_data))


model_accuracy <- sum(alc_data$high_use == predictions) / nrow(alc_data)


simple_guess_accuracy <- sum(alc_data$high_use == simple_guess_predictions) / nrow(alc_data)

cat("Model Accuracy:", model_accuracy, "\n")
cat("Simple Guessing Accuracy:", simple_guess_accuracy, "\n")

```

The model has 100% success rate and no false results. There are also 0 incorrectly typed individuals. 

#Bonus I

```{r}
install.packages("caret")
library(caret)

alc_data$high_use <- as.factor(alc_data$high_use)

set.seed(123)

ctrl <- trainControl(method = "cv", number = 10)

cv_model <- train(high_use ~ Dalc + Walc + romantic, data = alc_data, method = "glm", family = binomial, trControl = ctrl)

print(cv_model)

previous_model_error <- 0.26  

cv_model_error <- 1 - cv_model$results$Accuracy

cat("\nPrevious Model Error:", previous_model_error, "\n")
cat("Cross-Validation Model Error:", mean(cv_model_error), "\n")

```

Bonus I: Performing a 10 fold cross validation and checking the test performance. 

In the first section, the coding contains creation of the training control. Since both accuracy and Kappa are 1, this means the model has a perfect classification and agreement of outome respectively. It also performs better than the prevous model, since the error is 0.

#Bonus II
```{r}

```
