#Assignment 4

The dataset used for this weeks assignment describes the housing values in Suburbs of Boston. The set is shown below. The meaning of the factors are as follows:crim = per capita crime rate by town, zn = proportion of residential land zoned for lots over 25,000 sq.ft., indus = proportion of non-retail business acres per town, chas = Charles River dummy variable (= 1 if tract bounds river; 0 otherwise), nox = nitrogen oxides concentration (parts per 10 million), rm = average number of rooms per dwelling, age = proportion of owner-occupied units built prior to 1940, dis = weighted mean of distances to five Boston employment centres, rad = index of accessibility to radial highways, tax = full-value property-tax rate per $10,000, ptratio = pupil-teacher ratio by town, black = 1000(Bk−0.63)^2 where Bk is the proportion of blacks by town, lstat = lower status of the population (percent), medv = median value of owner-occupied homes in $1000s
```{r}
library(MASS)

data(Boston)
str(Boston)
dim(Boston)
```
Showing a graphical overview of the data and correlations. It is used to predict the value of owner-occupied homes.
```{r}
par(mar = c(2, 2, 1, 1))
par(mfrow = c(4, 4))
for (i in 1:14) {
  plot(Boston[, i], main = names(Boston)[i], col = "green", pch = 20)
}
par(mfrow = c(1, 1))
par(mar = c(5, 4, 4, 2) + 0.1) 
```
# standardizing the data set and creating a training set.
```{r}
scaled_data <- as.data.frame(scale(Boston))

summary(scaled_data)

scaled_data$crime_category <- cut(scaled_data$crim, breaks = quantile(scaled_data$crim), labels = FALSE)

scaled_data <- scaled_data[, -which(colnames(scaled_data) == "crim")]

set.seed(123)

train_indices <- sample(1:nrow(scaled_data), 0.8 * nrow(scaled_data))
train_set <- scaled_data[train_indices, ]
test_set <- scaled_data[-train_indices, ]
```
Mean of each variable is now close to 0 and standard deviation is close to 1. 

Linear discriminant analysis
```{r}

library(ggplot2)

lda_model <- lda(crime_category ~ ., data = train_set)

lda_data <- data.frame(predict(lda_model, train_set)$x, crime_category = train_set$crime_category)

ggplot(lda_data, aes(x = LD1, y = LD2, color = crime_category)) +
  geom_point() +
  ggtitle("LDA Biplot") +
  xlab("LD1") +
  ylab("LD2") +
  theme_minimal()
```

In the plot shown above the x and y axes represent the linear discriminant functions. They are linear combinations of the original variables.  
LD1: the first linear discriminant function maximizes the separation between different categories
LD2:The second linear discriminant function, which is orthogonal to the former, captures additional variation not explained by the LD1. 
In the plot we can crealy see that the higher crime categories (mainly 4) are separated from the rest. This means less violent crimes can be separated from the rest. 

#Predicting classes with he LDA model

```{r}
actual_crime_categories <- test_set$crime_category

test_set <- test_set[, -which(colnames(test_set) == "crime_category")]

predicted_classes <- predict(lda_model, newdata = test_set)$class

confusion_matrix <- table(actual = actual_crime_categories, predicted = predicted_classes)

print(confusion_matrix)

accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)
cat("Overall Accuracy:", round(accuracy, 4), "\n")
```
The plot above consists of tru positive, true negative, false positive and false negative values. Hence numbers 1-4. Within the plot we can se instances fitting these categories. The overall accuracy is (TP+TN)/Total. In this case the overall accuracy is 65% whis is not great. 

