---
output:
  html_document: default
  pdf_document: default
---

# Assignment 6

```{r}
(date)
```

```{r}

library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(MASS)
library(corrplot)
library(GGally)
library(tibble)
library(FactoMineR)
library(factoextra)
library(lme4)

rats <- read.table("data/rats.txt", sep = ",", header = T)

str(rats)

```

Changing the categorical variables ID, Group and WD
```{r}
rats$ID <- factor(rats$ID)
rats$Group <- factor(rats$Group)
rats$WD <- factor(rats$WD)
str(rats)

```
The data compares three different diet types (groups) and their effect on the rats development ie mass. The weight has been measured over 9 weeks with 11 different time points.
```{r}
rats$ID <- factor(rats$ID)
rats$Group <- factor(rats$Group)
rats$WD <- factor(rats$WD)
str(rats)

```
Weight development over time in different groups
```{r}

ggplot(rats, aes(x = time, y = weight, linetype=ID, col=Group)) +geom_line() + scale_linetype_manual(values = rep(1:10, times=4)) + theme(legend.position = "none") + scale_y_continuous(limits = c(min(rats$weight), max(rats$weight))) + facet_grid(. ~ Group, labeller = label_both) + xlab("time (days)")
```
Plotting values using standardized values. Shows the distribution better in the beginning.
```{r}
rats <- rats %>% group_by(time) %>% mutate(stdweight = (weight-mean(weight))/sd(weight)) %>% ungroup()

ggplot(rats, aes(x = time, y = stdweight, linetype=ID, col=Group)) +geom_line() + scale_linetype_manual(values = rep(1:10, times=4)) + theme(legend.position = "none") + facet_grid(. ~ Group, labeller = label_both) + xlab("time (days)") + ylab("scaled weight")
```
As we can see the Group 1 has a much lower weight to from the very beginning, it has also some minimal increase throughout the observation period. Groups 2 and 3 have higher weight increase throughout the observation period however 

```{r}
rats2 <- rats %>% group_by(Group, time) %>% summarise( mean = mean(weight), se = (sd(weight)/sqrt(length(weight)))) %>% ungroup()

#average weight change over the observation period
ggplot(rats2, aes(x = time, y = mean, linetype = Group, shape = Group, col=Group)) + geom_line() + scale_linetype_manual(values = c(1,2,3)) + geom_point(size=3) + scale_shape_manual(values = c(1,2,3)) + geom_errorbar(aes(ymin=mean-se, ymax=mean+se, linetype="1"), width=0.3) + scale_y_continuous(name = "mean(weight) +/- se(weight)") + scale_x_continuous(name = "time (days)")

#total change in weight
rats3 <- rats %>% filter(time > 1) %>% group_by(Group, ID) %>% summarise(mean=mean(weight)) %>% ungroup()
ggplot(rats3, aes(x = Group, y = mean, col=Group)) + geom_boxplot() + stat_summary(fun = "mean", geom = "point", shape=23, size=2.5, fill = "white") + scale_y_continuous(name = "mean(weight), days 8-64")

rats4 <- subset(rats3, rats3$mean<550)

ggplot(rats4, aes(x = Group, y = mean, col=Group)) + geom_boxplot() + stat_summary(fun = "mean", geom = "point", shape=23, size=2.5, fill = "white") + scale_y_continuous(name = "mean(weight), days 8-64")
```

Group 1 has the smallest weight. Group 2 has the biggest increase in weight gain. Group 3 have the biggest weight however the gain is not as big as in group 2. 

##T-test to evaluate the group differences and suitability into the linear model

```{r}
rats5 <- subset(rats4, rats4$Group==1 | rats4$Group==2)
t.test(mean ~ Group, data = rats5, var.equal=TRUE)

#Group 1 versus Group 2
```

```{r}
rats5 <- subset(rats4, rats4$Group==1 | rats4$Group==3)
t.test(mean ~ Group, data = rats5, var.equal=TRUE)

#Group 1 versus Group 3
```

```{r}
rats5 <- subset(rats4, rats4$Group==2 | rats4$Group==3)
t.test(mean ~ Group, data = rats5, var.equal=TRUE)

#Group 2 versus Group 3
```

```{r}
rats5 <- subset(rats4, rats4$Group==2 | rats4$Group==3)
t.test(mean ~ Group, data = rats5, var.equal=TRUE)

#Group 2 versus Group 3
```

```{r}
rats5 <- subset(rats3, rats3$Group==2 | rats4$Group==3)
t.test(mean ~ Group, data = rats5, var.equal=TRUE)

baseline <- subset(rats, rats$time==1)
rats5 <- rats3 %>% mutate(baseline = baseline$weight)

fit <- lm(mean ~ baseline + Group, data = rats5)

anova(fit)
```

Comparing Group 1 to Groups 2 and 3 has small p-values. Hence there is a significant difference in the weight. Comparing groups 2 and 3 shows already a smaller p -value. Moreover removing the outlines increases the p-value even more questioning the correlation.

Based on the Anova analysis it might be good to choose rats with a similar starting weight as that is the strongest predictive value. 

#BPRS

```{r}
bprs <- read.table("data/bprs.txt", header=T, sep = ",")

str(bprs)
```

```{r}
bprs$subject <- 1:40

bprs$treatment <- factor(bprs$treatment)
bprs$subject <- factor(bprs$subject)
bprs$week <- factor(bprs$week)
str(bprs)
```
```{r}

```
